SIG Proceedings Paper in LaTeX Format

### 2019-08-30 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/contribution.tex
structure change w/ freq words

\subsubsection{Analyzing word frequencies}
. For this matter we chose to analyze the distribution of word frequencies. A box plot of the distribution, on the text8 dataset, can be found in Figure \ref{fig:boxplot_freq}. 
\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{images/no_sampling_boxplot}
\caption{Distribution of word frequencies in the text8 datasedt}
\label{fig:boxplot_freq}
\end{figure} 
This shows that very few words appear very frequently, and are therefore responsible for the building of the most context pairs. For example the word \textit{the} creates 50\% of the pairs in the text8 dataset. In consequence if we create arbitrary batches, most words  will appear multiple times in a batch and the learning will be hindered. A solution would be to only allow words to appear once in a batch, in the next subsection we present such a solution with a variable batch size. 
\subsubsection{Criteria for batches}
To solve the above problem we only chose to block multiple appearance in a batch on the center words. The context words can be the same for two different center words.  Each batch will contain the highest possible number of pairs where each center word does not appear twice. As with all batches, each pair will only appear once in a batch. Therefore the batch size will be very high at the beginning of the training and will then decrease very rapidly. In consequence the average batch size is small and training time is increased, in comparison to arbitrary created batches. To remedy this problem we tried to reduce the number of very small batches by altering the dataset, while maintaining the quality of learning. We call it, deletion of outliers and explain it in the next subsection. 
\subsubsection{Deletion of outliers}
To solve the above matter, we looked  at the distribution of word frequencies (Figure \ref{fig:boxplot_freq}). We experimented with different settings (mean, median, average, outlier) and found that deleting all words that had a frequency greater than the 99 percentile was best. 
This technique drastically reduces the data set size, it also deletes words that appear very frequently, it therefore reduces the number of very small batches.

>>>>

\subsubsection{Analyzing word frequencies}
. For this matter we chose to analyze the distribution of word frequencies. A box plot of the distribution, on the text8 dataset, can be found in Figure \ref{fig:boxplot_freq}. 
\begin{figure}[h]
\centering
\includegraphics[scale=0.3]{images/no_sampling_boxplot}
\caption{Distribution of word frequencies in the text8 datasedt}
\label{fig:boxplot_freq}
\end{figure} 
This shows that very few words appear very frequently, and are therefore responsible for the building of the most context pairs. For example the word \textit{the} creates 50\% of the pairs in the text8\footnote{See Section \ref{sec:evalutaion}} dataset. In consequence if we create arbitrary batches, most words  will appear multiple times in a batch and the learning will be hindered. A solution would be to only allow words to appear once in a batch, in the next subsection we present such a solution with a variable batch size. 
\subsubsection{Criteria for batches}
To solve the above problem we only chose to block multiple appearance in a batch on the center words. The context words can be the same for two different center words.  Each batch will contain the highest possible number of pairs where each center word does not appear twice. As with all batches, each pair will only appear once in a batch. Therefore the batch size will be very high at the beginning of the training and will then decrease very rapidly. In consequence the average batch size is small and training time is increased, in comparison to arbitrary created batches where the batch size is fixed. To remedy this problem we tried to reduce the number of very small batches by altering the dataset, while maintaining the quality of learning. We call it, deletion of outliers and explain it in the next subsection. 
\subsubsection{Deletion of outliers}
To solve the above matter, we looked  at the distribution of word frequencies (Figure \ref{fig:boxplot_freq}). We experimented with different settings (mean, median, average, outlier) and found that deleting all words that had a frequency greater than the 99 percentile was best. 
This technique drastically reduces the data set size,by deleting words that appear very frequently, it therefore reduces the number of very small batches.

### 2019-08-30 SAC/submission/ACM_SigConf_SAC2020-2020_1.0/includes/results.tex
discussion added improvements word freq
structure change w/ freq words

\section{Results}\label{sec:results}
As stated in the previous section we ran multiple experiments for each optimizer with the above configuration and datasets. . This Section will  give an overview of the experimental results. Each subsection will give an explanation over the achieved result with a specific optimizer.

\subsection{Simple Batch Version}
\subsubsection{SGD}
The first challenge for each optimizer was to find a correct learning rate. As SGD is the optimizer used in Gensim \citep{gensim} we first tried the same learning rate as the default value in Gensim \citep{gensim}, i.e 0.01,  and then performed a random search to find a better one. As expected a bell curve shape resulted, a learning rate that is too high leads to diversion and a learning rate that is too low leads to a training time that is too slow. The best value that we found for the learning rate is $0.0075$. With this setting SGD converged in 11 epochs. The second experiment was to add input shuffling.
As seen in Figure \ref{fig:results_sgd}, for almost every learning rate the convergence time decreased. Our model, with the best setting, know converges in only 7 epochs. Another interesting fact to point out from Figure \ref{fig:results_sgd} is that with input shuffling we achieved better results with higher learning rates. As for learning rates of $0.01$ and $0.025$ we did converge in 11 epochs with input shuffling but did not converge in 20 epochs without it.

@@

@@

\begin{table}[tb]
    \caption{Results on the text8 dataset}
    \begin{tabular}{l r r }
        \toprule
Model    & Convergence Time & Word Similarity \\ \hline
SGD & {11}              & 0.65            \\ 
SGD w/shuffling & {7}              & 0.66            \\ 
Adam & {3}              & 0.66            \\ 
Adam w/ shuffling & {2}              & 0.66      \\ 
Gensim   & 4          & 0.66            \\ \hline
        %
\textbf{Adam w/o outliers }& \textbf{1} & 0.66 \\  \hline
\bottomrule
\end{tabular}
\end{table}

>>>>

\section{Results}\label{sec:results}
As stated in the previous section we ran multiple experiments for each optimizer with the above configuration and datasets . This Section will  give an overview of the experimental results. 
The Section is structured as follows: The first part will describe the results achieved by our batched version with advanced optimizers and input shuffling the second part will focus on batches with improved quality and the deletion of outliers.

\subsection{Simple Batch Version}
This subsection will give an explanation over the achieved results with different optimizers, combined with input shuffling on our batched version of the Skip-Gram Model with negative sampling (SGNS). It will first explain the result achieved with Stochastic Gradient Descent (SGD) followed by the result achieved with adam as optimizers. 
\subsubsection{SGD}
The first challenge for each optimizer was to find a correct learning rate. As SGD is the optimizer used in Gensim \citep{gensim} we first tried the same learning rate as the default value in Gensim \citep{gensim}, i.e 0.01,  and then performed a random search to find a better one. As expected a bell curve shape resulted, a learning rate that is too high leads to diversion and a learning rate that is too low leads to a training time that is too slow. The best value that we found for the learning rate is $0.0075$. With this setting SGD converged in 11 epochs. The second experiment was to add input shuffling.
As seen in Figure \ref{fig:results_sgd}, for almost every learning rate the convergence time decreased. Our model, with the best setting, know converges in only 7 epochs. Another interesting fact to point out from Figure \ref{fig:results_sgd} is that with input shuffling we achieved better results with higher learning rates. As for learning rates of $0.01$ and $0.025$ we did converge in 11 epochs with input shuffling but did not converge in 20 epochs without it.

@@

This conlcudes the results achieved with arbitrary batches. The following section will cover the results achieved when creating batches that only contains center words once and how the deletion of outliers affect the results.

@@

\begin{table}[tb]
    \caption{Results on the text8 dataset}
    \begin{tabular}{l r r }
        \toprule
Model    & Convergence Time & Word Similarity \\ \hline
SGD & {11}              & 0.65            \\ 
SGD w/shuffling & {7}              & 0.66            \\ 
Adam & {3}              & 0.66            \\ 
Adam w/ shuffling & {2}              & 0.66      \\ 
Gensim   & 4          & 0.66            \\ \hline
        %
\textbf{Adam w/o outliers }& \textbf{1} & 0.66 \\  \hline
\bottomrule
\end{tabular}\label{table:results_text8}
\end{table}
